---
output:
  html_document:
    theme: spacelab
    df_print: paged
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
params:
  proj: null
  t0: null
  eval_year: 2022
  input_dir: null
  output_dir: null
  fullname: null
  shapefile_path: null
  country_path: null
  pairs_path: null
  carbon_density_path: null
  additionality_path: null
  verbose: FALSE
  branch: null
---

```{r include=FALSE}

# TO KNIT THIS NOTEBOOK, RUN THE FOLLOWING LINE IN A POWERSHELL TERMINAL:

# Rscript -e "rmarkdown::render(input='evaluations/pipeline_results.Rmd', output_file='/maps/aew85/tmf_pipe_out/example_project.html', params=list(proj='example_project', t0=2010, ...))"

# Mandatory args: proj, t0, eval year
# Optional args: input dir, output dir, fullname, country_path, shapefile path, pairs_path, carbon density path, verbose
# You must either specify input dir and output dir OR provide absolute paths to each of the objects required (pairs, shapefile, carbon density, additionality, country list)
# Verbose option includes additional descriptive text. Defaults to false.
# Include name of pipeline branch with branch parameter for reproducibility.

```

```{r settings, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, warning=FALSE,message=FALSE)

library(tidyverse)
library(sf)
library(reshape2)
library(maps)
library(mapdata)
library(ggspatial)
library(arrow)
library(rnaturalearth)
library(rnaturalearthdata)
library(stringr)
library(jsonlite)
library(countrycode)
library(here)

```

```{r inputs, echo=FALSE,warning=FALSE, message=FALSE}

# Extract params

project_name <- params$proj
start_year <- as.numeric(params$t0)
evaluation_year <- as.numeric(params$eval_year)
verbose <- params$verbose

```

---
title: "`r paste0('4C Ex-Post Evaluation: ', if (!is.null(params$fullname)) { params$fullname } else { gsub('_', ' ', project_name) %>% stringr::str_to_title() })`"
subtitle: "`r format(Sys.Date(), "%B %Y")`"
---

```{r set_paths, echo=FALSE,warning=FALSE, message=FALSE}

# get script path

script_path <- here('evaluations/scripts') # change as appropriate

# get data path

if (!is.null(params$output_dir)) {
  data_path <- paste0(params$output_dir,'/',project_name)
}

# get input path

if (!is.null(params$input_dir)) {
  input_dir <- params$input_dir 
}

# get branch

if (!is.null(params$branch)) {
  branch <- params$branch
}

```

```{r shapefile_path, echo=FALSE, message=FALSE}

# add error message for shapefile

if (is.null(params$input_dir) && is.null(params$shapefile_path)) {
    warning("Error: insufficient information to read shapefile. To map the shapefile, you must provide either input_dir OR shapefile_path in params.")}

# read shapefile

if (!is.null(params$shapefile_path)) {
  shapefile_path <- params$shapefile_path
} else if(file.exists(input_dir)) { shapefile_path <- paste0(input_dir,'/',project_name,'.geojson') }
if(file.exists(shapefile_path)) {shapefile <- read_sf(shapefile_path)}

```

```{r pairs_path, echo=FALSE, message=FALSE}

# add error message for pairs

if (is.null(params$output_dir) && is.null(params$pairs_path)) {
    warning("Error: insufficient information to read pairs. To analyse pairs, you must provide either output_dir OR pairs_path in params.")}

# get path to pairs

if (!is.null(params$pairs_path)) {
  pairs_path <- params$pairs_path
} else if(file.exists(data_path)) {pairs_path <- file.path(data_path,'pairs') }

```

```{r carbon_density_path, echo=FALSE, message=FALSE}

# add error message for carbon density

if (is.null(params$output_dir) && is.null(params$carbon_density_path)) {
    }

# read carbon density

if (!is.null(params$carbon_density_path)) {
  carbon_density_path <- params$carbon_density_path
} else if(file.exists(data_path)) {carbon_density_path <- list.files(data_path,pattern='carbon',full.names=T)[1] }

if(file.exists(carbon_density_path)) {
    carbon_density <- read.csv(carbon_density_path)
    } else {
        warning("Error: insufficient information to read carbon density. To print carbon density information, you must provide either output_dir OR carbon_density_path in params.")}

```

```{r country_path, echo=FALSE, message=FALSE}

# add error message for country

if (is.null(params$output_dir) && is.null(params$country_path)) {
    warning("Error: insufficient information to read country. To print country information, must provide either output_dir OR country_path in params.")}

# read country path

if (!is.null(params$country_path)) {
  country_path <- params$country_path
} else if(file.exists(data_path)) {country_path <- list.files(path = data_path, pattern = "country-list", full.names = TRUE)[1]}

```

```{r additionality_path, echo=FALSE, message=FALSE}

# add error message for additionality

if (is.null(params$output_dir) && is.null(params$additionality_path)) {
    warning("Error: insufficient information to read additionality. To print additionality information, you must provide either output_dir OR additionality_path in params.")}

if (!is.null(params$additionality_path)) {
  additionality_path <- params$additionality_path
} else if(file.exists(data_path)) {additionality_path <- list.files(path = data_path, pattern = "additionality", full.names = TRUE)[1]}
if(file.exists(additionality_path)) {additionality <- read.csv(additionality_path)}

```

```{r read_pairs, echo=FALSE}
  
# get filenames and filter for matched points

files_full_raw <- list.files(pairs_path,
                          pattern='*.parquet',full.names=T,recursive=F)
files_full <- files_full_raw[!grepl('matchless',files_full_raw)]
files_short_raw <- list.files(path=pairs_path,
                          pattern='*.parquet',full.names=F,recursive=F)
files_short <- files_short_raw[!grepl('matchless',files_short_raw)]

# initialise dfs

vars <- c(colnames(read_parquet(files_full[1])),'pair')
paired_data_raw <- data.frame(matrix(ncol=length(vars),nrow=0))
colnames(paired_data_raw) <- vars
paired_data_raw$pair <- as.factor(paired_data_raw$pair)

for(j in 1:length(files_full)){
  
  # read in all parquet files for a given project
  
  f <- data.frame(read_parquet(files_full[j]))
  
  # add identity column
  
  f$pair <- as.factor(c(replicate(nrow(f),files_short[j])))
  
  # append data to bottom of df
  
  paired_data_raw <- bind_rows(paired_data_raw,f)
  
}

# generate separate datasets for project and counterfactual 

project <- paired_data_raw %>% dplyr::select(starts_with('k'),pair)
cf <- paired_data_raw %>% dplyr::select(starts_with('s'),pair)

# create project-counterfactual merged dataset

colnames(cf) <- colnames(project)
pair_merged <- bind_rows(project,cf)
names(pair_merged) <- str_sub(names(pair_merged),3)
names(pair_merged)[names(pair_merged) == "ir"] <- "pair"
data <- pair_merged %>% select_if(~ !any(is.na(.)))

# merge reference data to cf-counterfactual merged dataset with type column

type <- c(replicate(nrow(project),'Project'),replicate(nrow(cf),'Counterfactual'))
data$type <- type

```

```{r get_shapefile_area, echo=FALSE}

if(exists("shapefile")){
    project_area_ha <- as.numeric(sum(st_area(st_make_valid(shapefile)))/10000)
}
  
```

```{r get_country_names}

if(exists("country_path")){

    # define function for extracting country names 

    get_country_names <- function(country_codes_path) {
        codes <- as.character(fromJSON(country_codes_path))
        country_names <- countrycode(codes, 'iso2c', 'country.name')
        country_names[country_names == 'Congo - Kinshasa'] <- 'Democratic Republic of the Congo'
        return(country_names)
    }

    # read in country json and get names
    country_vec <- get_country_names(country_path)

    # define function for printing the country names if there are multiple

    if (length(country_vec) > 1) {
        country_string <- paste(country_vec[-length(country_vec)], collapse = ", ")
        country_string <- paste(country_string, "and", country_vec[length(country_vec)])
        } else {
        country_string <- country_vec[1]
    }

}


```

\

# About the project

`r project_name %>% str_replace_all("_", " ") %>% str_to_title()` is located in `r if(exists("country_string")) country_string`. The project started in `r start_year` and has an area of `if (exists("project_area_ha")) r format(project_area_ha, big.mark = ",", scientific = FALSE, digits = 3)` hectares.

```{r echo=FALSE}

# ________________ FILL IN PROJECT-SPECIFIC INFORMATION __________________

# Replace this chunk with a short narrative about the context of the project and why it was chosen for evaluation by 4C. Include any other details relevant to the interpretation of this document.

```



\

# Introduction to the 4C methodology

`r if(verbose){"

The 4C method for calculating the additionality of a project is based on pixel matching. This approach allows us to identify places which are similar to the project but are not protected. We can then measure deforestation in these places and use this as our counterfactual scenario (what would have happened in the absence of the project) against which we measure the impact that the project has had.

More information about 4C's approach can be found below.

[Quantify Earth, an interactive platform which demonstrates our pixel-matching approach in action](https://quantify.earth/#/globe/1)

[4C explainer page](https://4c.cst.cam.ac.uk/about/additionality-leakage-and-permanence)

[Cambridge Carbon Impact: an evaluation of some existing carbon projects](https://www.cambridge.org/engage/coe/article-details/6409c345cc600523a3e778ae)

[The full PACT methodology](https://www.cambridge.org/engage/coe/article-details/657c8b819138d23161bb055f)

"}`



\

# Additionality summary

The graph below shows the annual trend in additionality from 10 years before the project start to the present day. The solid grey vertical line represents the project start, whilst the dashed grey horizontal line represents 0 additionality (i.e. no avoided deforestation). Above this line, the project has experienced less forest carbon loss than the counterfactual; below this line, it has experienced more forest carbon loss than the counterfactual.

\

```{r additionality_summary, echo=FALSE}

if(exists("additionality")) {
    additionality %>% ggplot(aes(x = year, y = additionality)) +
    geom_vline(xintercept = start_year, alpha = 0.4) +
    geom_hline(yintercept = 0, alpha = 0.4, linetype = 'dashed') +
    #annotate(geom='text',x=start_year,y=10000,label='Project start',size=5,colour='grey50')+
    geom_line() +
    xlab('Year') +
    ylab(expression(paste('Additionality (Mg ',  CO[2], "e)", sep = '')))+
    theme_classic()
} else {print("Additionality information not available")}



```

\

Raw values are also presented below.

\

```{r echo=FALSE}

if(exists("additionality")) {
    additionality %>% rename(Additionality = additionality, Year = year)
} else {print("Additionality information not available")}

```

\

# Detailed explanation of results

The following sections will detail how we arrived at the additionality results, including the location and quality of the matched points, the deforestation rates in each set, and the carbon density values used to convert these deforestation rates to carbon emissions reductions.

\

### Location of matched points

We sampled `r format((nrow(data)/2), big.mark = ",", scientific = FALSE, digits = 3)` points within the project and an equal number of matching points from outside of the project. These matching points serve as our counterfactual scenario for deforestation.

Below we show the location of the counterfactual matching points (shown in blue) relative to the project (shown in red), both at the country and project scale.

`r if(nrow(data) > 20000){"Note that, for clarity and computational efficiency, we show only 10% of the points."}`

\

```{r match_locations, warning=FALSE, echo=FALSE, message=FALSE}

if(exists("shapefile") & exists("country_vec") & exists("data")) {

    # downsample by 90%

    if(nrow(data) > 20000){
    data_forplot <- data %>% sample_frac(0.1)
    } else {
    data_forplot <- data
    }

    # plot location of matching points

    country_map <- ne_countries(country = country_vec, returnclass = "sf", scale= "large")

    # transform crs

    shapefile <- st_transform(shapefile, st_crs(country_map))

    ggplot(data=country_map) + 
    geom_sf(colour=NA,fill='grey80')+
    geom_point(data=data_forplot,mapping=aes(x=lng,y=lat,colour=type),alpha=0.5) + 
    geom_sf(data=shapefile,fill='red',colour=NA,inherit.aes=F)+
    scale_color_manual(values=c('blue','red'))+
    coord_sf()+
    theme_void()+
    annotation_scale(text_cex=1.5,location='tl')+
    theme(legend.title = element_blank(),
            text=element_text(size=20))

    xmin <- filter(data, type=='Project') %>% select(lng) %>% min()
    xmax <- filter(data, type=='Project') %>% select(lng) %>% max()
    ymin <- filter(data, type=='Project') %>% select(lat) %>% min()
    ymax <- filter(data, type=='Project') %>% select(lat) %>% max()

    ggplot(data=country_map) + 
    geom_sf(colour=NA,fill='grey80')+
    geom_point(data=data_forplot,mapping=aes(x=lng,y=lat,colour=type),alpha=0.5) + 
    geom_sf(data=shapefile,fill='red',colour=NA,inherit.aes=F)+
    scale_color_manual(values=c('blue','red'))+
    coord_sf(xlim=c(xmin-0.5,xmax+0.5),ylim=c(ymin-0.5,ymax+0.5))+
    theme_void()+
    annotation_scale(text_cex=1.5,location='tl')+
    theme(legend.title = element_blank(),
            text=element_text(size=16),
            legend.position='none')

} else {print("Insufficient information available to map points")}


```

### Quality of matches

Here we show how well the matching points align with the project in terms of our matching variables. Correspondence between the project (shown in red in the plots below) and the counterfactual (shown in blue) indicates that the counterfactual will faithfully represent the business-as-usual scenario for places like the project.

-   Inaccessibility (motorized travel time to healthcare, minutes)

-   Slope ($^\circ$)

-   Elevation (meters)

-   Forest cover at t0 (project start, %)

-   Deforestation at t0 (%)

-   Forest cover at t-5 (5 years prior to project start, %)

-   Deforestation at t-5 (%)

-   Forest cover at t-10 (10 years prior to project start, %)

-   Deforestation at t-10 (%)

Forest cover and deforestation are measured as the proportion of pixels within a 1km radius of a particular point which are classified either as undisturbed forest (in the case of forest cover) or deforested (in the case of deforestation) by the JRC Tropical Moist Forest dataset.

More information about the datasets we use can be found below:

[MAP access to healthcare](https://malariaatlas.org/project-resources/accessibility-to-healthcare/)

[SRTM elevation data](https://www.earthdata.nasa.gov/sensors/srtm)

[JRC tropical moist forest dataset](https://forobs.jrc.ec.europa.eu/TMF)

\

```{r match_quality, warning=FALSE,message=FALSE, echo=FALSE}

if(exists("data")) {

    # plot matches

    source(file.path(script_path,'plot_matchingvars.R'))

    plot_matching_variables(data)

} else {print("Insufficient information available to evaluate match quality")}



```

\

### Standardised mean differences

We can quantify the similarity in matching variables in terms of their standardised mean difference (SMD). An SMD of \< 0.25 indicates that points are well-matched for that particular variable.

In the below plot, the blue points indicate the SMD value (i.e. the amount of difference between the project and counterfactual) for each variable. The grey dotted lines at (-0.25, +0.25) represent the boundary within which our differences would ideally fall in order for the project and counterfactual to be considered well-matched.

\

```{r smd}

if(exists("data")) {

    source(file.path(script_path,'std_mean_diff.R'))

    results <- std_mean_diff(pairs_path)

    # changing sign for interpretation

    results$smd <- (-1)*results$smd

    # changing order of variables

    variables <- c('cpc10_d','cpc5_d','cpc0_d',
                    'cpc10_u','cpc5_u','cpc0_u',
                    'access','slope','elevation')

    # plotting

    ggplot(results,aes(x=smd,y=variable))+
    #geom_boxplot(outlier.shape=NA,colour='blue')+
    geom_point(colour='blue',fill='blue',alpha=0.3,size=4)+
    geom_vline(xintercept=0)+
    geom_vline(xintercept=0.25,lty=2,colour='grey30')+
    geom_vline(xintercept=-0.25,lty=2,colour='grey30')+
    scale_y_discrete(labels=c(bquote(Deforestation~t[-10]~("%")),
                                bquote(Deforestation~t[-5]~("%")),
                                bquote(Deforestation~t[0]~("%")),
                                bquote(Forest~cover~t[-10]~("%")),
                                bquote(Forest~cover~t[-5]~("%")),
                                bquote(Forest~cover~t[0]~("%")),
                                'Inaccessibility (mins)',paste0('Slope (',intToUtf8(176),')'),'Elevation (m)'))+
    xlab('Standardised mean difference')+
    xlim(-1,1)+
    theme_classic()+
    theme(axis.title.y=element_blank(),
            legend.title=element_blank(),
            legend.box.background=element_rect(),
            legend.position='none',
            text=element_text(size=20),
            axis.text.y=element_text(size=20))

} else {print("Insufficient information available to evaluate match quality")}

```

\

### Deforestation within the project

Now focusing on deforestation within the project, we can examine the spatial distribution of the following 4 processes pertinent to forest carbon stock changes:

-   Undisturbed forest to degraded forest 

-   Degraded forest to deforested land 

-   Undisturbed forest to deforested land 

-   Undisturbed land to reforested land

\

These transitions are shown in the plot below for the `r evaluation_year-start_year`-year period between `r start_year` and `r evaluation_year`. They are overlaid on the project area, shown in grey. If a transition is not shown, it did not occur in the period examined.

Data from [JRC tropical moist forest dataset](https://forobs.jrc.ec.europa.eu/TMF).

\

```{r deforestation_spatial_distribution, echo=FALSE, warning=FALSE}


if(exists("data")) {

    # plot deforestation within project

    source(file.path(script_path,'plot_transitions.R'))

    plot_transitions(data=data,t0=start_year,period_length=evaluation_year-start_year,shapefile=shapefile)

} else {print("Insufficient information available to evaluate deforestation")}


```

\

### Deforestation and degradation rates within project and counterfactual

Here we compare various deforestation processes between the project and counterfactual. We present average annual rates measured between `r start_year` and `r evaluation_year`.

\

```{r proportions_undisturbed_degraded, echo=FALSE}

if(exists("data")) {

# obtaining the area of undisturbed and degraded forest at t0, for use later

source(file.path(script_path,'def_rate.R'))

proj_data <- data  %>% filter(type=='Project')

prop_und <- get_prop_class(data=proj_data,t0=start_year-10,class=1)
prop_deg <- get_prop_class(data=proj_data,t0=start_year-10,class=2)

} else {print("Insufficient information available to evaluate deforestation")}

```

***Rate of forest loss, %/year***

First we can calculate the average annual rate at which undisturbed forest is lost. This refers to the % loss of undisturbed tropical moist forest per year, i.e. it is relative to the amount of tropical moist forest present at the beginning of the project.

```{r rate_of_forest_loss_percent, echo=FALSE}

if(exists("data")) {

    source(file.path(script_path,'def_rate.R'))
    
    df <- def_rate(data=data,t0=start_year,period_length=evaluation_year-start_year)

    df %>% t() %>% data.frame() %>% rename('Rate of forest loss (%/year)' = 1)

} else {print("Insufficient information available to evaluate deforestation")}


```

\

***Separate deforestation and degradation processes, %/year***

The rate of forest loss can be broken down into more specific processes, presented in the table below:

-   degradation of undisturbed forest

-   deforestation of undisturbed forest

-   deforestation of degraded forest

-   reforestation of undisturbed forest


\

```{r separate_deforestation_processes_percent, echo=FALSE}

if(exists("data")) {

    source(file.path(script_path,'def_rate.R'))
    
    df_sep <- def_rate_seperate(data=data,t0=start_year,period_length=evaluation_year-start_year)

    df_sep

} else {print("Insufficient information available to evaluate deforestation")}

```

\

We can also convert these deforestation rates to hectares per using the following formula:

\

$$ {\text{Deforestation rate (hectares/year)}} =
\left( \frac{\text{Deforestation rate (%/year)}}{100} \right) \times \text{Project area (hectares)} \times \text{Proportion of forest type present at } t_0
$$

\

It is necessary to correct for the amount of each forest type (undisturbed or degraded) present at the beginning of the project. For this project these proportions are as follows:

-   Undisturbed forest: `r format(100*prop_und, digits = 3)`%

-   Degraded forest: `r format(100*prop_deg, digits = 3)`%

The rates of overall forest loss and individual deforestation processes are shown in hectares in the tables below.

\

***Rate of forest loss, hectares/year***

```{r rate_of_forest_loss_ha, echo=FALSE}

if(exists("data")) {

    df_ha <- df

    df_ha[1,1:2] <- (df_ha[1,1:2]/100)*project_area_ha*prop_und

    colnames(df_ha) <- c('Project','Counterfactual')

    df_ha %>% t() %>% data.frame() %>% rename('Rate of forest loss (ha/year)' = 1)

} else {print("Insufficient information available to evaluate deforestation")}

```

\

***Separate deforestation and degradation processes, hectares/year***

```{r separate_deforestation_processes_ha, echo=FALSE}

if(exists("data")) {

    source(file.path(script_path,'def_rate.R'))

    df_sep_ha <- df_sep

    df_sep_ha[df_sep_ha$`Forest type`=='Undisturbed forest',4] <- (df_sep_ha[df_sep_ha$`Forest type`=='Undisturbed forest',4]/100)*project_area_ha*prop_und

    df_sep_ha[df_sep_ha$`Forest type`=='Disturbed forest',4] <- (df_sep_ha[df_sep_ha$`Forest type`=='Disturbed forest',4]/100)*project_area_ha*prop_deg

    df_sep_ha %>% rename('Rate (ha/year)' = 4)

} else {print("Insufficient information available to evaluate deforestation")}

```

\

***Land cover changes over time***

Presenting the above data in another way, we can visualise the year-on-year change in different land cover classes (undisturbed forest, degraded forest, deforested land and regrowth) for both the project and the counterfactual.

In the below plots, the vertical grey dashed line represents the start year of the project.

```{r luc_timeseries_all, echo=FALSE}

if(exists("data")) {

    source(file.path(script_path,'land_cover_timeseries.R'))

    df <- get_luc_timeseries(data,t0=start_year,tend=evaluation_year)

    df %>% mutate(
    luc = as.factor(luc),
    year = as.numeric(year)) %>% 
    ggplot(aes(x=year,y=percentage,colour=luc,lty=type))+
    geom_line(linewidth=1.5,alpha=0.5)+
    geom_vline(xintercept=start_year,lty=2,colour='grey30')+
    scale_colour_manual(values=c('darkgreen','gold2','orange3','steelblue2'),
                        name='Land Use Class',labels=c('Undisturbed forest','Degraded forest','Deforested land','Regrowth'))+
    scale_linetype_manual(name='Location',values=c('solid','dotted'),breaks=c('Project','Counterfactual'))+
    xlab('Year')+
    ylab('% cover')+
    theme_classic()

} else {print("Insufficient information available to evaluate deforestation")}

```

\

***Land cover changes over time: undisturbed forest only***

Zooming in on the trend in **undisturbed forest** cover for project and counterfactual (shown in green in the plot above), we expect the trends in forest cover to be parallel prior to the project start (indicating that they are well-matched) but diverging after the project start, indicating a measurable reduction in deforestation under the project scenario.

Here the trajectories are shown in red (project) and blue (counterfactual). They show the mean and 95% confidence intervals of % undisturbed tropical moist forest cover, calculated across the 100 sets of points we generate as part of our bootstrapping procedure. More details about this algorithm are available in the [PACT Methodology](https://www.cambridge.org/engage/api-gateway/coe/assets/orp/resource/item/647a14a14f8b1884b7b97b55/original/pact-tropical-moist-forest-accreditation-methodology.pdf).

\


```{r luc_timeseries, echo=FALSE}

if(exists("data")) {

    source(file.path(script_path,'land_cover_timeseries.R'))

    # caching result of land cover time series
    result <- luc_class1_uncertainty(data=data, t0=start_year, tend=evaluation_year)

    # calculating stats + plotting

    result %>%
    group_by(type,year) %>%
    summarise(mean=mean(percent_class1),
                se = sd(percent_class1) / sqrt(n()), # Standard error
                t_critical = qt(0.975, df = n() - 1), # Critical t-value for 95% CI
                ci_lower = mean - t_critical * se,
                ci_upper = mean + t_critical * se,
                .groups = 'drop') %>% 
    ggplot(aes(x = year, y = mean, color = type)) +
    geom_line(size = 1) + # Line for the mean
    geom_ribbon(aes(ymin = ci_lower, ymax = ci_upper, fill = type), alpha = 0.2) + # Confidence interval ribbon
    labs(
        x = "Year",
        y = "% undisturbed forest") +
    theme_classic() +
    geom_vline(xintercept=start_year,lty=2,colour='grey30')+
    scale_color_manual(values = c("Project" = "red", "Counterfactual" = "blue")) + 
    scale_fill_manual(values = c("Project" = "red", "Counterfactual" = "blue")) +
    theme(legend.title = element_blank())

} else {print("Insufficient information available to evaluate deforestation")}

```

\

### Carbon densities

In order to calculate additionality, the deforestation rates are converted to carbon emissions rates using regional carbon density values generated through NASA GEDI data. These are presented in the table below for each land use class, each of which is associated with a different carbon density value. 

More information on GEDI data is available [here](https://www.earthdata.nasa.gov/sensors/gedi).

\

```{r carbon_density, echo=FALSE}

if(exists("carbon_density")) {

    carbon_density <- carbon_density %>% mutate(
    land.use.class = case_when(
        land.use.class == 1 ~ 'Undisturbed',
        land.use.class == 2 ~ 'Degraded',
        land.use.class == 3 ~ 'Deforested',
        land.use.class == 4 ~ 'Reforested',
        land.use.class == 5 ~ 'Water',
        land.use.class == 6 ~ 'Other')
    )


    colnames(carbon_density) <- c('Land Use Class', 'Carbon Density (Mg/Ha)')

    carbon_density

} else {print("Carbon density not available")}

```

\

The additionality summary presented at the top of this document is based on the difference in the carbon loss estimates between the project and the counterfactual scenario.

\

`r if(verbose){"


# Statement on leakage and permanence

Leakage and permanence are two factors that affect the long-term emissions reductions contributed by a project but **have not been included in this evaluation**.

**Leakage** is the displacement of activities which deplete forest carbon stocks from the project to other areas due to the implementation of the project. In the worst case scenario, 100% of these activities are displaced, effectively nullifying the additionality of the project. Leakage can be reduced by interventions which remove the incentive to continue activities which deplete forest carbon stocks in areas outside of the project.

**Permanence** is the ability of a project to protect carbon stocks long-term. Carbon stored in forests is inherently impermanent, given the finite lifespan of trees and the potential for deforestation and catastrophic events such as wildfires. The estimates given in this evaluation assume that all carbon stored is permanent, but in reality this is unlikely to be the case.

\

You can find out more about our plans to deal with leakage and permanence in our [explainer page](https://4c.cst.cam.ac.uk/about/additionality-leakage-and-permanence), and in [the full PACT methodology](https://www.cambridge.org/engage/api-gateway/coe/assets/orp/resource/item/647a14a14f8b1884b7b97b55/original/pact-tropical-moist-forest-accreditation-methodology.pdf).

"}`

---

This report was generated on `r format(Sys.Date(), "%B %d, %Y")` using the `r branch` of the [tmf-implementation code](https://github.com/quantifyearth/tmf-implementation).
