---
output:
  html_document:
    theme: spacelab
    df_print: paged
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
params:
  proj: null
  t0: null
  input_dir: null
  output_dir: null
  fullname: null
  country_path: null
  shapefile_path: null
  pairs_path: null
  carbon_density_path: null
  branch: null
---

```{r include=FALSE}

# TO KNIT THIS NOTEBOOK, RUN THE FOLLOWING LINE IN A SHELL TERMINAL:

# Rscript -e "rmarkdown::render(input='~/evaluations/R/ex_ante_evaluation_template.Rmd', output_file='~/evaluations/R/example_project.html', params=list(proj='example_project', t0=2010, ...))"

# Mandatory args: proj, t0
# Optional args: input dir, output dir, fullname, country_path, shapefile path, pairs_path, carbon density path
# You must either specify input dir and output dir OR provide absolute paths to each of the objects required

```

```{r settings, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, warning=FALSE,message=FALSE)

library(tidyverse)
library(sf)
library(reshape2)
library(maps)
library(mapdata)
library(ggspatial)
library(arrow)
library(rnaturalearth)
library(rnaturalearthdata)
library(rnaturalearthhires)
library(stringr)
library(jsonlite)
library(countrycode)
library(scales)
library(here)
library(patchwork)
library(knitr)
library(kableExtra)

```

```{r read_inputs, echo=FALSE,warning=FALSE, message=FALSE}

project_name <- params$proj
start_year <- as.numeric(params$t0)
branch <- params$branch

```

---
title: "`r paste0('4C Ex-Ante Evaluation: ', if (!is.null(params$fullname)) { params$fullname } else { gsub('_', ' ', project_name) %>% stringr::str_to_title() })`"
subtitle: "`r format(Sys.Date(), "%B %Y")`"
---

```{r set_paths, echo=FALSE,warning=FALSE, message=FALSE}

# get output format

output_format <- ifelse(knitr::is_latex_output(), "latex", "html")

# get script path

script_path <- here('scripts')

# get explainer diagram path 

diagram_path <- here('methods_diagram.png')

# get data path

if (!is.null(params$output_dir)) {
  data_path <- paste0(params$output_dir,'/',project_name)
}

# get path to pairs

if (!is.null(params$pairs_path)) {
  pairs_path <- params$pairs_path
} else { pairs_path <- file.path(data_path,'pairs') }

# read shapefile

if (!is.null(params$input_dir)) {
  input_dir <- params$input_dir 
}

if (!is.null(params$shapefile_path)) {
  shapefile_path <- params$shapefile_path
} else { shapefile_path <- paste0(input_dir,'/',project_name,'.geojson') }
shapefile <- read_sf(shapefile_path)

# read carbon density

if (!is.null(params$carbon_density_path)) {
  carbon_density_path <- params$carbon_density_path
} else { carbon_density_path <- list.files(data_path,pattern='carbon',full.names=T)[1] }
carbon_density <- read.csv(carbon_density_path)

# read country path

if (!is.null(params$country_path)) {
  country_path <- params$country_path
} else { country_path <- list.files(path = data_path, pattern = "country-list", full.names = TRUE)}

```

```{r read_pairs, echo=FALSE}
  
# get filenames and filter for matched points

files_full_raw <- list.files(pairs_path,
                          pattern='*.parquet',full.names=T,recursive=F)
files_full <- files_full_raw[!grepl('matchless',files_full_raw)]
files_short_raw <- list.files(path=pairs_path,
                          pattern='*.parquet',full.names=F,recursive=F)
files_short <- files_short_raw[!grepl('matchless',files_short_raw)]

# initialise dfs

vars <- c(colnames(read_parquet(files_full[1])),'pair')
paired_data_raw <- data.frame(matrix(ncol = length(vars), nrow = 0)) %>%
  setNames(vars) %>%
  mutate(
    pair = as.factor(pair),
    k_trt = as.factor(k_trt),
    s_trt = as.factor(s_trt)
  )

for(j in 1:length(files_full)){
  
  # read parquet file
  
  f <- data.frame(read_parquet(files_full[j]),check.names = FALSE)
  
  # add identity column
  
  f$pair <- as.factor(c(replicate(nrow(f),str_remove(files_short[j], "\\.parquet$"))))
  
  # append data to bottom of df
  
  paired_data_raw <- bind_rows(paired_data_raw,f)
  
}

# generate separate datasets for project and counterfactual 

project <- paired_data_raw %>% dplyr::select(starts_with('k'),pair)
cf <- paired_data_raw %>% dplyr::select(starts_with('s'),pair)

# create project-counterfactual merged dataset

colnames(cf) <- colnames(project)
pair_merged <- bind_rows(project,cf)
names(pair_merged) <- str_sub(names(pair_merged),3)
names(pair_merged)[names(pair_merged) == "ir"] <- "pair"

# add type column and remove excess cols

data <- pair_merged %>%
        mutate(type=c(replicate(nrow(project),'Project'),replicate(nrow(cf),'Counterfactual'))) %>% 
        select(-c(contains('trt'),ID))

```

```{r get_shapefile_area, echo=FALSE}
  
project_area_ha <- as.numeric(sum(st_area(st_make_valid(shapefile)))/10000)

```

```{r get_country_names}

# define function for extracting country names 

get_country_names <- function(country_codes_path) {
    codes <- as.character(fromJSON(country_codes_path))
    country_names <- countrycode(codes, 'iso2c', 'country.name')
    country_names[country_names == 'Congo - Kinshasa'] <- 'Democratic Republic of the Congo'
    return(country_names)
  }

# get country names

country_vec <- get_country_names(country_path)

  # define function for printing the country names if there are multiple

  if (length(country_vec) > 1) {
    country_string <- paste(country_vec[-length(country_vec)], collapse = ", ")
    country_string <- paste(country_string, "and", country_vec[length(country_vec)])
  } else {
    country_string <- country_vec[1]
  }


```

\

# Introduction

This Report has been prepared by researchers at the Cambridge Centre for Carbon Credits (4C) and has been funded by a charitable grant from the Tezos Foundation.  4C utilises innovative, evidence-based approaches to examine the scientific basis of nature-based carbon conservation initiatives and, insodoing, provides a way for different stakeholders to assess the quality of carbon credits (ex post and/or ex ante).  

**Disclaimer:  Nothing in this Report constitutes formal advice or recommendations, an endorsement of proposed action, or intention to collaborate; instead, it sets out the details of an evaluation using a method which is still under development. The Report is considered complete as of the publication date shown, though methods are likely to change in future.**

\

# About the project

`r if (!is.null(params$fullname)) { params$fullname } else { gsub('_', ' ', project_name) %>% stringr::str_to_title() }` is located in `r country_string`. The proposed area measures `r format(project_area_ha, big.mark = ",", scientific = FALSE, digits = 3)` hectares. 

For the purposes of this evaluation, we have set the proposed start date to `r start_year`.

```{r echo=FALSE}

# ________________ FILL IN PROJECT-SPECIFIC INFORMATION __________________

# Replace this chunk with a short narrative about the context of the project and why it was chosen for evaluation by 4C. Include any other details relevant to the interpretation of this document.

```



\

# Introduction to the 4C method

*Our method for forecasting ex-ante additionality remains under development.*

The 4C approach to forecasting additionality involves identifying places that experienced similar deforestation levels in the past as the project area does today. We start by analyzing forest cover changes in the project area between 10 years ago and the present day. Using pixel-matching techniques, we then identify comparable places outside the project that experienced similar deforestation trends between 20 and 10 years ago (the *matching period*). This allows us to match the deforestation trajectory of the project with that of the matched pixels, but offset in time. This concept is illustrated by the left-hand diagonal arrow in the figure below.

We can consider the matched pixels as a historical representation of the project as it is today. By examining deforestation in the matched pixels over the subsequent 10 years (the *baseline period*), we estimate a *baseline prediction* — the deforestation expected in the project area under the counterfactual (business-as-usual) scenario. This rate is then projected forward over the next 10 years, as illustrated by the right-hand diagonal arrow in the figure below. We convert the deforestation rate to carbon dioxide emissions using best estimates of carbon density.

```{r, echo=FALSE, fig.align='center', fig.width=6}

knitr::include_graphics(diagram_path)

```


Making predictions about future deforestation is challenging, and there are multiple sources of uncertainty at play. These include: the quantification of carbon, the choice of matching pixels, the effect of leakage and impermanence, future political changes and market forces. We are constantly improving our method in order to minimise these uncertainties. Due to the inherent uncertainty associated with ex-ante (before-the-fact) predictions, carbon credits should only ever be quantified and issued ex-post (after the fact).

More information about 4C's approach to impact evaluation can be found below.

[Quantify Earth, an interactive platform which demonstrates our pixel-matching approach in action](https://quantify.earth/#/globe/1)

[4C explainer page](https://4c.cst.cam.ac.uk/about/additionality-leakage-and-permanence)

[Cambridge Carbon Impact: an evaluation of some existing carbon projects](https://www.cambridge.org/engage/coe/article-details/6409c345cc600523a3e778ae)

[Our paper on the social value of impermanent carbon credits](https://www.nature.com/articles/s41558-023-01815-0)

[The PACT methodology for ex-post evaluations](https://www.cambridge.org/engage/coe/article-details/657c8b819138d23161bb055f)

\


# Methods

The following sections will detail how we arrived at a forecast of future deforestation and the potential to generate additionality by reducing this deforestation. This includes the location and quality of the matched points, the deforestation rates in each set of points, and the carbon density values used to convert these deforestation rates to carbon emissions reductions.

\

### Location of matched points

We sampled `r format((nrow(data)/2), big.mark = ",", scientific = FALSE, digits = 3)` points within the project and an equal number of matching points from outside of the project. We used these matched points to make a prediction of the counterfactual scenario for deforestation.

Below we show the location of the matching points (shown in blue) relative to the project (shown in red), both at the country and project scale.

`r if(nrow(data) > 20000){"Note that, for clarity and computational efficiency, we show only 10% of the points."}`

```{r match_locations, warning=FALSE, echo=FALSE, message=FALSE}

# downsample no. of points by 90%

if(nrow(data) > 20000){
  data_forplot <- data %>% sample_frac(0.1)
} else {
  data_forplot <- data
}

# plot location of matching points

country_map <- ne_countries(country = country_vec, returnclass = "sf", scale= "large")

# transform crs

shapefile <- st_transform(shapefile, st_crs(country_map))

ggplot(data=country_map) + 
  geom_sf(colour=NA,fill='grey80')+
  geom_point(data=data_forplot,mapping=aes(x=lng,y=lat,colour=type),alpha=0.5,size=0.5) + 
  scale_color_manual(values=c('blue','red'),labels=c('Matched points','Project'))+
  coord_sf()+
  theme_void()+
  annotation_scale(text_cex=1.5,location='bl')+
  theme(legend.title = element_blank(),
        text=element_text(size=16),
        legend.position='none')

xmin <- filter(data, type=='Project') %>% select(lng) %>% min()
xmax <- filter(data, type=='Project') %>% select(lng) %>% max()
ymin <- filter(data, type=='Project') %>% select(lat) %>% min()
ymax <- filter(data, type=='Project') %>% select(lat) %>% max()

ggplot(data=country_map) + 
  geom_sf(colour=NA,fill='grey80')+
  geom_point(data=data_forplot,mapping=aes(x=lng,y=lat,colour=type),alpha=0.5,size=0.5) + 
  scale_color_manual(values=c('blue','red'),labels=c('Matched points','Project'))+
  coord_sf(xlim=c(xmin-0.5,xmax+0.5),ylim=c(ymin-0.5,ymax+0.5))+
  theme_void()+
  annotation_scale(text_cex=1.5,location='bl')+
  theme(legend.title = element_blank(),
        text=element_text(size=16),
        legend.position='none')

```

### Quality of matches

Here we show how well the matching points align with the project in terms of our matching variables. Correspondence between the project (shown in red in the plots below) and the matched points (shown in blue) indicates that the the matched points are composed of places that are similar to the project in terms of the drivers of deforestation and are expected to exhibit similar deforestation trends.

-   Inaccessibility (motorized travel time to healthcare, minutes)

-   Slope ($^\circ$)

-   Elevation (meters)

-   Forest cover at t0 (start year, %)

-   Deforestation at t0 (%)

-   Forest cover at t-5 (5 years prior to start year, %)

-   Deforestation at t-5 (%)

-   Forest cover at t-10 (10 years prior to start year, %)

-   Deforestation at t-10 (%)

Forest cover and deforestation are measured as the proportion of pixels within a 1km radius of a particular point which are classified either as undisturbed forest (in the case of forest cover) or deforested (in the case of deforestation) by the JRC Tropical Moist Forest dataset.

More information about the datasets we use can be found below:

[MAP access to healthcare](https://malariaatlas.org/project-resources/accessibility-to-healthcare/)

[SRTM elevation data](https://www.earthdata.nasa.gov/sensors/srtm)

[JRC tropical moist forest dataset](https://forobs.jrc.ec.europa.eu/TMF)

\

```{r match_quality, warning=FALSE,message=FALSE, echo=FALSE}

# plot matches

source(file.path(script_path,'plot_matchingvars.R'))

plot_matching_variables(data,ex_ante='true')

```

\

### Standardised mean differences

We can quantify the similarity in matching variables in terms of their standardised mean difference (SMD). The SMD allows us to quantify the similarity between the project and the matched points in a way that is comparable across variables. 

In the below plot, the blue points indicate the SMD value (i.e. the amount of difference between the project and matched points, in standard deviations) for each variable. Values further from zero indicate a larger difference. The grey dotted lines at (-0.25, +0.25) represent the boundary within which our SMDs would ideally fall in order for the project and matched points to be considered well-matched.

\

```{r smd}

std_mean_diff <- function(pairs_path) {
  
  # clean data

  files_full_raw <- list.files(pairs_path,
                               pattern='*.parquet',full.names=T,recursive=F)
  files_full <- files_full_raw[!grepl('matchless',files_full_raw)]
  files_short_raw <- list.files(path=pairs_path,
                                pattern='*.parquet',full.names=F,recursive=F)
  files_short <- files_short_raw[!grepl('matchless',files_short_raw)]
  
  # initialise dfs
  
  vars <- c(colnames(read_parquet(files_full[1])),'pair')
  df <- data.frame(matrix(ncol=length(vars),nrow=0))  %>% 
  setNames(vars) %>% 
  mutate(k_trt=as.factor(k_trt),
         s_trt=as.factor(s_trt))
  
  for(j in 1:length(files_full)){
    
    # read in all parquet files for a given project
    
    f <- data.frame(read_parquet(files_full[j])) %>% 
         mutate(k_trt=as.factor(k_trt),
                s_trt=as.factor(s_trt))
    
    # append data to bottom of df
    
    df <- bind_rows(df,f)
    
  }

  # calculate smd

  smd_results <- data.frame(variable = character(), smd = numeric(), stringsAsFactors = FALSE)
  
  variables <- c('cpc10_d','cpc5_d','cpc0_d',
                 'cpc10_u','cpc5_u','cpc0_u',
                 'access','slope','elevation')
    
    for (var in variables) {
      k_var <- df[[paste0("k_", var)]]
      s_var <- df[[paste0("s_", var)]]
      
      k_mean <- mean(k_var, na.rm = TRUE)
      s_mean <- mean(s_var, na.rm = TRUE)
      k_sd <- sd(k_var, na.rm = TRUE)
      s_sd <- sd(s_var, na.rm = TRUE)
      
      pooled_sd <- sqrt((k_sd^2 + s_sd^2) / 2)
      smd <- (k_mean - s_mean) / pooled_sd
      
      smd_results <- rbind(smd_results, data.frame(variable = var, smd = smd, stringsAsFactors = FALSE))
    }
    
  return(smd_results)
}

results <- std_mean_diff(pairs_path)

# changing sign for interpretation

results$smd <- (-1)*results$smd

# changing order of variables

variables <- c('cpc10_d','cpc5_d','cpc0_d',
                 'cpc10_u','cpc5_u','cpc0_u',
                 'access','slope','elevation')

results$variable <- factor(results$variable, levels=variables)

# plotting

 ggplot(results,aes(x=smd,y=variable))+
  #geom_boxplot(outlier.shape=NA,colour='blue')+
  geom_point(colour='blue',fill='blue',alpha=0.3,size=4)+
  geom_vline(xintercept=0)+
  geom_vline(xintercept=0.25,lty=2,colour='grey30')+
  geom_vline(xintercept=-0.25,lty=2,colour='grey30')+
  scale_y_discrete(labels=c(bquote(Deforestation~t[-10]~("%")),
                            bquote(Deforestation~t[-5]~("%")),
                            bquote(Deforestation~t[0]~("%")),
                            bquote(Forest~cover~t[-10]~("%")),
                            bquote(Forest~cover~t[-5]~("%")),
                            bquote(Forest~cover~t[0]~("%")),
                            'Inaccessibility (mins)',paste0('Slope (',intToUtf8(176),')'),'Elevation (m)'))+
  xlab('Standardised mean difference')+
  xlim(-1,1)+
  theme_classic()+
  theme(axis.title.y=element_blank(),
        legend.title=element_blank(),
        legend.box.background=element_rect(),
        legend.position='none',
        text=element_text(size=14),
        axis.text.y=element_text(size=14))


```

\

### Deforestation within the project

Now focusing on deforestation within the project, we can examine the spatial distribution of the following 4 processes pertinent to forest carbon stock changes:

-   Undisturbed forest to degraded forest 

-   Degraded forest to deforested land 

-   Undisturbed forest to deforested land 

-   Undisturbed land to reforested land (which indicates that regrowth occurred after a deforestation event)

\

These transitions are shown in the plot below for the 10-year period between `r start_year-10` and `r start_year`. They are overlaid on the project area which is shown in grey. If a transition is not shown, it did not occur in the period examined.

Data from [JRC tropical moist forest dataset](https://forobs.jrc.ec.europa.eu/TMF).

\

```{r deforestation_spatial_distribution, echo=FALSE, warning=FALSE}

# plot deforestation within project

source(file.path(script_path,'plot_transitions.R'))

proj_coords <- data %>% 
  filter(type=='Project') %>% 
  select(lat,lng)

proj_input_defplot <- data %>% 
  filter(type=='Project') %>% 
  select(contains('luc')) %>% 
  setNames(paste0("luc_", (start_year-10):start_year)) %>%
  cbind(proj_coords)

proj_input_defplot <- proj_input_defplot[, !is.na(colnames(proj_input_defplot))]

plot_transitions(data=proj_input_defplot,t0=start_year-10,period_length=10,shapefile=shapefile)

```

\

### Land cover changes within project and matched pixels

In the below plots, we show the changes in land classes over time for both the project (red) and the matched points (blue).

Note the following:

-   The vertical grey dashed line represents the start year of the project (`r start_year`). The timings shown on the x-axis are relative to this start year.

-   As explained in the 'Methods' section, the matched points are offset in time relative to the project by 10 years. This means that all changes observed in the matched points happened 10 years prior to the equivalent time point in the project. This time offset allows us to use the last 10 years in the matched points as a prediction of the next 10 years for the project.

-   Solid lines represent ex-post observed changes, whereas the dotted line represents the prediction for the future of the project. 

```{r make_inputs, echo=FALSE}

# preparing inputs

proj_input <- data %>% 
  filter(type=='Project') %>% 
  select(contains('luc')) %>% 
  setNames(paste0("luc_", (start_year-10):start_year))
proj_input <- proj_input[, !is.na(colnames(proj_input))]


cf_input <- data %>% 
  filter(type=='Counterfactual') %>% 
  select(contains('luc')) %>% 
  setNames(paste0("luc_", (start_year-20):(start_year))) %>%
  select(where(~ all(!is.na(.))))

```

```{r luc_timeseries_all, echo=FALSE}

source(file.path(script_path,'land_cover_timeseries.R'))

# getting results

proj_results <- get_luc_timeseries(proj_input,t0=start_year-10,tend=start_year,type='single') %>% 
  mutate(type='Project')

cf_results <- get_luc_timeseries(cf_input,t0=start_year-20,tend=start_year,type='single') %>% 
  mutate(type='Counterfactual')

# combining results

results <- bind_rows(proj_results, cf_results)

```

Showing the trend for undisturbed, degraded, deforested and regrowth in turn:

```{r undisturbed_timeseries, fig.width=8,fig.height=13}

# add prediction from the matched pixels:

prediction <- cf_results %>% 
                filter(year >= (start_year-10))  %>% 
                mutate(type='Project',
                year=year+10)

results <- bind_rows(results,prediction)

# make a custom function for plotting the results

plot_timeseries <- function(luc_value, title_str) {
  
  #remove gap between solid and dotted project line
  percent_val <- results %>%
  filter(year == start_year 
          & type == "Project"
          & luc == luc_value) %>%
  pull(percentage)

  # df wrangling
  extended_results <- results %>%
    mutate(
      luc = as.numeric(luc),
      year = as.numeric(year),
      line_type = ifelse(type == "Project" & year > start_year, "dotted", "solid"),
      type = case_when(
      type == "Counterfactual" ~ "Matched points",
      TRUE ~ type
    )
    ) %>%
    bind_rows(data.frame(
      year = start_year,
      luc = luc_value,   
      percentage = percent_val,
      type = 'Project',
      line_type = 'dotted'
    ))

   extended_results %>%
    filter(luc == luc_value) %>%
    ggplot(aes(x = year, y = percentage, color = type, linetype = line_type)) +
    geom_line(linewidth = 1.5) +
    geom_vline(xintercept = start_year, linetype = 2, color = 'grey30') +
    #geom_vline(xintercept = start_year-10, linetype = 2, color = 'grey30') +
    scale_colour_manual(name = 'Location',
                        values = c('red','blue'),
                        breaks = c('Project', 'Matched points'),
                        labels = c('Project', 'Matched points'))+
    xlab('Year') +
    ylab('% cover') +
    ggtitle(title_str) +
    guides(linetype = "none") +
    theme_classic() +
    scale_linetype_manual(values = c("solid" = "solid", "dotted" = "dotted"))+
    facet_wrap(~type)+
    xlim(start_year-20,start_year+10)

}

plot_1 <- plot_timeseries(luc_value=1, title_str='Undisturbed forest') + theme(legend.position='none',axis.title.x = element_blank())
plot_2 <- plot_timeseries(luc_value=2, title_str='Degraded forest') + theme(legend.position='none', axis.title.x = element_blank())
plot_3 <- plot_timeseries(luc_value=3, title_str='Deforested land') + theme(legend.position='none', axis.title.x = element_blank())
plot_4 <- plot_timeseries(luc_value=4, title_str='Regrowth') + theme(legend.position='none', axis.title.x = element_text(size=14))

plot_1 + plot_2 + plot_3 + plot_4 + plot_layout(ncol=1)

```

### Deforestation rates in the matched points during the baseline period

```{r proportions_undisturbed_degraded, echo=FALSE}

# obtaining the area of undisturbed and degraded forest at t0, for use later

source(file.path(script_path,'def_rate.R'))

prop_und <- get_prop_class(data=proj_input,t0=start_year-10,class=1)
prop_deg <- get_prop_class(data=proj_input,t0=start_year-10,class=2)

```

Here we present the deforestation rates observed in the matched pixels over the past 10 years (the baseline period). 

Forest loss transitions can be broken down into the following processes: 

-   degradation of undisturbed forest

-   deforestation of undisturbed forest

-   deforestation of degraded forest

-   regrowth of undisturbed forest (implies previous deforestation)

We calculate the rate at which these processes occur in the matched pixels using the following method:

1. Calculate the percentage of matched pixels which have undergone one of the above processes (according to the JRC classification) during the baseline period. 
2. Divide this percentage by 10 to give an annual rate of change, in %/year, relative to the amount of (undisturbed or degraded) forest present at the beginning of the project.
3. Multiply this rate by the area of the (undisturbed or degraded) forest  10 years prior to the project start to give an annual rate, in hectares per year.

The amounts of forest in the project area 10 years prior to project start are as follows:

-   Undisturbed forest: `r format(100*prop_und, big.mark = ",", scientific = FALSE, digits = 3)`%

-   Degraded forest: `r format(100*prop_deg, big.mark = ",", scientific = FALSE, digits = 3)`%

The rates are given below.

```{r rate_of_forest_loss_ha, echo=FALSE}

source(file.path(script_path,'def_rate.R'))

df_rate_percent <- def_rate_single(data=cf_input,t0=start_year-10,period_length=10)

df_rate_ha <- df_rate_percent

df_rate_ha[df_rate_ha$`Forest type`=='Undisturbed forest',3] <- (df_rate_ha[df_rate_ha$`Forest type`=='Undisturbed forest',3]/100)*project_area_ha*prop_und

df_rate_ha[df_rate_ha$`Forest type`=='Disturbed forest',3] <- (df_rate_ha[df_rate_ha$`Forest type`=='Disturbed forest',3]/100)*project_area_ha*prop_deg

knitr::kable(
  df_rate_ha %>% 
  rename('Rate (ha/year)' = 3) %>%
  mutate(across(where(is.numeric), ~ sprintf("%.2f", .)))
)


```

\

### Carbon stock changes in the matched points during the baseline period

Here we present the carbon density calculations for this project.

In order to convert land cover changes to carbon emissions, we use regional aboveground (AGB) carbon density values generated through NASA GEDI data. 

More information on GEDI data is available [here](https://www.earthdata.nasa.gov/sensors/gedi).

Note that, in calculating carbon stock changes, we assume the following:

- Belowground biomass (BGB) is 20% of AGB (based on Cairns et al. 1997)

- Deadwod biomass is 11% of AGB (based on IPCC 2003)

- Emitted carbon dioxide is 47% of total biomass (based on Martin and Thomas 2011)


\
```{r additionality_forecast}

baseline_stocks <- data.frame(matrix(nrow=2*nrow(carbon_density),ncol=8))
colnames(baseline_stocks) <- c('time','luc','agb','bgb','dwb','total_c','area','total_byarea')
luc_counter <- 1
row_counter <- 1

carbon_density <- filter(carbon_density, land.use.class %in% c(1:6))

for(i in carbon_density$land.use.class){

  for(j in c("Start","End")) {

    # get agb

    agb <- carbon_density$carbon.density[luc_counter]

    # get other values

    bgb <- agb*0.2
    dw <- agb*0.11
    total <- agb + bgb + dw
    #total_co2 <- total*0.47 # we're doing this step later

    # get area of class i

    if (j == "Start") {
      area_of_forest <- get_prop_class(cf_input,t0=start_year-10,class=i)*project_area_ha
    } else if (j == "End") {
      area_of_forest <- get_prop_class(cf_input,t0=start_year,class=i)*project_area_ha }

    # multiply total by area
    
    total_byarea <- total*area_of_forest

    # adding to df

    baseline_stocks[row_counter,1] <- j
    baseline_stocks[row_counter,2] <- i
    baseline_stocks[row_counter,3] <- agb
    baseline_stocks[row_counter,4] <- bgb
    baseline_stocks[row_counter,5] <- dw
    baseline_stocks[row_counter,6] <- total
    baseline_stocks[row_counter,7] <- area_of_forest
    baseline_stocks[row_counter,8] <- total_byarea

    row_counter <- row_counter+1

  }

  # advance counter

  luc_counter <- luc_counter + 1

}

# formatting bits

baseline_stocks_format <- baseline_stocks
baseline_stocks_format <- baseline_stocks_format %>% filter(time == 'Start')
baseline_stocks_format <- baseline_stocks_format[2:6]

colnames(baseline_stocks_format) <- c(
                                 'Land use class',
                                 'AGB density (t C / ha)',
                                 'BGB density (t C / ha)',
                                 'Deadwood biomass density (t C / ha)', 
                                 'Total biomass density (t C / ha)', 
                                 'Total biomass (t C)')


# renaming classes

baseline_stocks_format <- baseline_stocks_format %>%
  mutate(`Land use class` = case_when(
    `Land use class` == "1" ~ 'Undisturbed',
    `Land use class` == "2" ~ 'Degraded',
    `Land use class` == "3" ~ 'Deforested',
    `Land use class` == "4" ~ 'Reforested',
    `Land use class` == "5" ~ 'Water',
    `Land use class` == "6" ~ 'Other',
    TRUE ~ as.character(`Land use class`)  # ensure no unexpected values
  ))


baseline_stocks_format[2:6] <- lapply(baseline_stocks_format[, 2:6], function(x) {
  if (is.numeric(x)) comma(x) else x
})

# Print only carbon calculations at this stage

baseline_stocks_format %>% 
  drop_na() %>% 
  mutate(across(where(is.numeric), ~ sprintf("%.2f", .))) %>% 
  kable(escape = FALSE, format = ifelse(knitr::is_latex_output(), "latex", "html")) %>%
  kable_styling(bootstrap_options = "striped")

```

# Results: baseline rate of carbon emissions

In this section we present the annual rate of carbon loss due to deforestation in the matched points during the baseline period. We can take this to be a prediction of the counterfactual scenario for the project (the *baseline*). 

First we present the carbon stock changes observed in the matched points during the baseline period:

```{r results}

baseline_stock_changes <- baseline_stocks[c(1:2,7:8)]

# reshape

reshaped_data <- baseline_stock_changes %>%
  mutate(luc = as.character(luc)) %>%
  group_by(luc) %>%
  summarize(
    area_start = area[time == "Start"],
    area_end = area[time == "End"],
    area_diff = area_start - area_end,
    c_start = total_byarea[time == "Start"],
    c_end = total_byarea[time == "End"],
    c_diff = c_start - c_end,
    .groups = 'drop'
  )

# get totals

total_row <- reshaped_data %>%
  summarize(
    luc = "Total",
    area_start = sum(area_start, na.rm = TRUE),
    area_end = sum(area_end, na.rm = TRUE),
    area_diff = sum(area_diff, na.rm = TRUE),
    c_start = sum(c_start, na.rm = TRUE),
    c_end = sum(c_end, na.rm = TRUE),
    c_diff = sum(c_diff, na.rm = TRUE)
  ) %>% 
  mutate(luc = as.character(luc))

baseline_stock_changes <- bind_rows(reshaped_data, total_row)

# add in conversion to CO2

baseline_stock_changes <- baseline_stock_changes  %>% 
    mutate(co2_diff = 0.47*c_diff)

# formatting bits

baseline_stock_changes_format <- baseline_stock_changes %>%
  mutate(across(where(is.numeric), ~ comma(.))) %>%
  mutate(across(where(is.numeric), ~ sprintf("%.2f", as.numeric(.))))

if (knitr::is_html_output()) {
  colnames(baseline_stock_changes_format) <- c(
                                 'Land use class',
                                 'Area at start (ha)',
                                 'Area at end (ha)',
                                 'Area loss (ha)',
                                 'Biomass at start (t)',
                                 'Biomass at end (t)',
                                 'Biomass loss (t)'
                                 'CO<sub>2</sub> loss (t)')
} else if (knitr::is_latex_output()) {
  colnames(baseline_stock_changes_format) <- c(
                                 'Land use class',
                                 'Area at start (ha)',
                                 'Area at end (ha)',
                                 'Area loss (ha)',
                                 'Biomass at start (t)',
                                 'Biomass at end (t)',
                                 'Biomass loss (t)'
                                 'CO$_{2}$ loss (t)')
}

baseline_stock_changes_format <- baseline_stock_changes_format %>%
  mutate(`Land use class` = case_when(
    `Land use class` == "1" ~ 'Undisturbed',
    `Land use class` == "2" ~ 'Degraded',
    `Land use class` == "3" ~ 'Deforested',
    `Land use class` == "4" ~ 'Reforested',
    `Land use class` == "5" ~ 'Water',
    `Land use class` == "6" ~ 'Other',
    TRUE ~ as.character(`Land use class`)  # ensure no unexpected values
  ))

baseline_stock_changes_format[nrow(baseline_stock_changes_format), 1] <- 'Total'

filtered_data <- baseline_stock_changes_format %>%
  drop_na() %>%
  mutate(across(where(is.numeric), ~ sprintf("%.2f", .)))

last_row_index <- nrow(filtered_data)

filtered_data %>%
  kable(escape = FALSE, format = ifelse(knitr::is_latex_output(), "latex", "html")) %>%
  kable_styling(bootstrap_options = "striped") %>%
  row_spec(last_row_index, bold = TRUE) 

```

```{r results_summary}

# find the difference

delta_c <- as.numeric(baseline_stock_changes[nrow(baseline_stock_changes), ncol(baseline_stock_changes)])
delta_c_annual <- delta_c/10

```

To calculate the baseline annual rate of carbon emissions, we sum the the differences in carbon stocks between the start and end of the baseline period, then divide the total by the length of the baseline period (10 years).

**For this project, the baseline annual rate of carbon emissions, in tonnes of carbon dioxide per year, is `r format(delta_c_annual, big.mark = ",", scientific = FALSE, digits = 3)`.** This should be interpreted as the maximum estimated number of carbon credits that could be generated if the project mitigated 100% of the baseline level of deforestation, assuming this is confirmed by ex post observations. We present alternative mitigation scenarios below.

### Expected additionality under different mitigation scenarios

Additionality depends not only on baseline deforestation rate but the ability of the project to mitigate that deforestation. Therefore, we present the expected additionality under different mitigation scenarios (100% to 10% mitigation success). The 100% mitigation scenario occurs where the project is able to mitigate all of the deforestation that would have occurred under the counterfactual scenario. This scenario is unlikely to be realistic, but gives a sense of the deforestation risk faced by each area. It is more likely that only a proportion of this baseline deforestation is mitigated - for example, under the 50% mitigation scenario, the project is able to reduce the deforestation rate by half. The higher the proportion of the baseline deforestation that is mitigated, the greater the additionality of a project.

Note that we present the raw additionality, without accounting for leakage and impermenance (discussed later).

We are in the process of producing confidence intervals that reflect the uncertainty associated with the baseline, which will be added to future revisions of this document.

```{r}

scenarios <- data.frame(matrix(ncol=2,nrow=5))
scenarios[1] <- c("10%","25%","50%","75%","100%")
scenarios[2] <- delta_c_annual*c(0.1,0.25,0.5,0.75,1)

if (knitr::is_html_output()) {
  colnames(scenarios) <- c('Scenario', 
                            'Additionality (t CO<sub>2</sub> / year)')
} else if (knitr::is_latex_output()) {
  colnames(scenarios) <- c('Scenario', 
                            'Additionality (t CO$_{2}$ / year)')
}

scenarios <- scenarios %>% 
  mutate(across(where(is.numeric), comma))

knitr::kable(
  scenarios
)

```

\

# Accounting for leakage and impermanence

Leakage and permanence are two factors that affect the long-term emissions reductions contributed by a project but **have not been included in this evaluation**.

**Leakage** is the displacement of activities which deplete forest carbon stocks from the project to other areas due to the implementation of the project. In the worst case scenario, 100% of these activities are displaced, effectively nullifying the additionality of the project. Leakage is likely to be lower if the processes leading deforestation and degradation do not result in high yielding land uses, or if the carbon densities within the project are high compared with those in other areas where these activities are taking place. Leakage can also be reduced by interventions which improve yields in areas already under production. We can provide guidance on how this could be achieved.

**Impermanence** occurs when the additionality generated by a project is reversed. Additional carbon stocks in forests are inherently vulnerable to these reversals. The estimates given in this evaluation assume that all carbon stored is permanent, but in reality this is unlikely to be the case. In future revisions of this document we aim to include indicative estimates of the equivalent permanence (the relative value of a impermanent credit relative to a permanent credit) for this project.

You can find out more about our plans to deal with leakage and permanence in our [explainer page](https://4c.cst.cam.ac.uk/about/additionality-leakage-and-permanence).

---

### Reproducibility

This report was generated on `r format(Sys.Date(), "%B %d, %Y")` using the `r branch` of the [tmf-implementation code](https://github.com/quantifyearth/tmf-implementation).
